---
title: "Beyond Exclusion:<br/>The Role of High-Stake Testing on Attendance"
font_family: 'Jost'
main_fontfamily: 'Yanone Kaffeesatz' 
author:
  - name: Magdalena Bennett
    affil: "1<br>"
    twitter: maibennett
    email: m.bennett@austin.utexas.edu
    main: true
  - name: Christopher Neilson
    affil: 2
  - name: Nicolás Rojas
    affil: 3
affiliation:
  - num: 1
    address: McCombs School of Business, The University of Texas at Austin
  - num: 2
    address: Economics Department, Princeton University 
  - num: 3
    address: Teachers College, Columbia University
main_findings: 
  - "Not only **low-performers attend less** on the day of the test in lower grades, but **high-performers attend more**"
  - '![](https://raw.githubusercontent.com/maibennett/presentations/main/content/presentations/Attendance/Salem_20220405/images/event_study_plot2nd-1.svg){.main_pic}'
  - "Using machine learning methods we can also **identify schools** more likely to **incentivize low-attendance** on bottom performers"
  - '![](https://raw.githubusercontent.com/maibennett/presentations/main/content/presentations/Attendance/Salem_20220405/mbennett_attendance_files/figure-html/prediction_example2-1.svg){.main_pic}'
main_picwidth: "90%"
logoleft_name: '![](https://raw.githubusercontent.com/maibennett/presentations/main/content/presentations/Attendance/Salem_20220405/images/qr_code_smallv2.svg){.main-img-left}'
primary_colour: "#35567D"
secondary_colour: "#6C5B7B"
accent_colour: "#F67280"
poster_height: "30in"
poster_width: "40in"
main_width: 0.4
title_textsize: "120pt"
main_textsize: "71pt"
body_textsize: "29pt"
output: 
  posterdown::posterdown_betterland:
    css: ["style.css"]
    self_contained: false
    pandoc_args: --mathjax
    highlight: haddock
    number_sections: false
    beforeInit: "macros.js"
link-citations: true
bibliography: packages.bib
---

```{r, include=FALSE}
knitr::opts_chunk$set(results = 'asis',
                      echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
```

## Introduction

**Non-representative patterns of attendance** can skew how useful test scores measures are for accomplishing their goal. The **main objectives** of this paper are the following:

```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

1. Understand the **average effect of testing on school attendance** across grades and performance 

2. **Identify schools** that **incentivize non-representative patterns** of attendance by **combining causal inference methods and machine learning**

3. Help improve current **imputation methods**


## Methods

### Event Study

$$Y_{ipsgt} = \sum_{P=1}^5\sum_{T=-4}^5 \tau^{PT}D^{PTG^*}_{ipsgt} + \gamma_{pt} +\alpha_i + \epsilon_{ipsgt}$$

- $Y_{ipsgt}$: Attendance (1,0) for student $i$, from GPA group $p$, in school $s$ and grade $g$ for day $t$.


- $D^{PTG^*}_{ipsgt}$: Indicator variable where $G^*$ is the tested grade.

### Prediction of counterfactual attendance

- **XGBoost** for large student panel of daily attendance
  - Includes FE by day of the week, school, grade, and student. Also includes sibling's attendance (if any) and attendance lag.
<br>

<li class="large-padding">Identify types of schools by clustering on *(Obs Attendance - Predicted Attendance)* using **K-means**.</li>

- Test different **imputation policies** and its consequences.

## Results

Students skip school on the day of the test. In lower grades, **lower-performers** attend less and **higher-performers** attend more, compared to a regular day. In higher grades, we **only observe action at the top of the distribution**

```{r event_study_plot_all, fig.height=6, fig.width=9, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(tidyverse)
library(hrbrthemes)
library(firasans)
library(haven)
library(scales)
library(patchwork)

# Directories
dir_data <- "C:/Users/mc72574/Dropbox/UT/UT Research/Asistencia/archivos_mai_18032022/"
dir_output <- "C:/Users/mc72574/Dropbox/Hugo/Sites/presentations/content/presentations/Attendance/Salem_20220405/images/"
dir_poster <- "C:/Users/mc72574/Dropbox/Hugo/Sites/presentations/content/presentations/Attendance/ACIC_20220524/"

# Color palettes:
text_color <- "#333f48"
#cols <- c("#0D0887","#5601A4","#900DA4","#BF3984","#E16462","#F89441","#FCCE25")
#cols <- c("#003f5c","#58508d","#bc5090","#ff6361","#ffa600")
cols <- c("#5A3D66","#9442B5","#4493AE","#A5AC20","#DAE058")

library(lemon)

temp1 <- haven::read_dta(paste0(dir_data, "temp_t3.dta"))
round_esp <- function(var){
  trunc(round(var, digits = 2)*100)/100
}
# Data to build the graph 
data_graph <- temp1 %>%
  dplyr::select(- stakes) %>% 
  filter(agno == 2014) %>%  
  pivot_longer(.,  starts_with("est") | starts_with("se") | starts_with("pval"), names_to = "variable", values_to = "value") %>% 
  mutate(variable = as.factor(variable)) %>% 
  mutate(variable = recode_factor(variable, `est1` = "D1", `est2` = "D2", `est3` = "D3D8", `est4` = "D9", `est5` = "D10", `est98` = "D10-D1", `est97` = "All")) %>% 
  filter(str_detect(variable, "se") == F & str_detect(variable, "pval") == F) %>%   
  mutate(value = value*100)  

gr <- data_graph  %>% 
  filter(variable != "All" & variable != "D10-D1") %>% 
  mutate(grado = as.factor(grado) %>% 
           forcats::fct_relevel("6", "8", "10", "2", "4") %>% 
           recode_factor(`6` = "6th Grade", `8` = "8th Grade", `10` = "10th Grade", `2` = "2nd Grade", `4` = "4th Grade"),
         variable =  variable %>% recode_factor(`D1` = "1st Decile", `D2` = "2nd Decile", `D3D8` = "3rd-8th Decile", `D9` = "9th decile", `D10` = "10th decile")) %>% 
  ggplot(., aes(x = variable, y = value)) +
  geom_segment( aes(x = variable, xend = variable, y = 0, yend = value), color = "grey") + 
    geom_hline(yintercept = 0, color = "grey") + 
  geom_hline(yintercept = 0, color = "grey") +  
  geom_hline(yintercept = -4, color = "grey90") +
  geom_hline(yintercept = -6, color = "grey90") + 
  geom_hline(yintercept = 4, color = "grey90") +  
  geom_hline(yintercept = 6, color = "grey90") +
  geom_hline(yintercept = -2, color = "grey90") +
  geom_hline(yintercept = 2, color = "grey90") + 
  geom_point( aes(x = variable, y = value, color = variable, fill=variable), pch = 21, size = 4, stroke = 1.1) +
  #theme_light() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  xlab("\n GPA Performance") +
  ylab("Change in Attendance \n") + theme_ipsum_fsc() + 
  scale_y_continuous(limits= c(-6, 4), breaks =  seq(from = -6, to = 4, by = 2), labels = dollar_format(suffix = "%", prefix = "")) + 
  facet_wrap(~grado, ncol = 3, as.table = F) + 
  scale_colour_discrete("GPA") + 
  theme(legend.title.align=0.5) + 
  scale_x_discrete(labels=c("1st Decile" = "D1", "2nd Decile" = "D2", "3rd-8th Decile" = "D3D8",
                            "9th decile" = "D9", "10th decile" = "D10")) + 
  scale_color_manual("GPA", values = c(cols[1], cols[2], cols[3], cols[4], cols[5])) + 
  scale_fill_manual("GPA", values = c(alpha(cols[1],0.8), alpha(cols[2],0.8), alpha(cols[3],0.8), 
                                      alpha(cols[4],0.8), alpha(cols[5],0.8))) + 
  theme_ipsum_fsc(plot_title_face = "bold",plot_title_size = 20) +
  labs(caption = "Note: p < 0.05 for all estimates except those touching the 0 bar. \n Markers symbols are the coefficients of the effect of testing on attendance.", color = text_color) +
      theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(),
          axis.line = element_line(colour = "dark grey")) +
  theme(axis.text=element_text(size = 16), axis.title = element_text(size = 16,face="bold")) + 
  theme(legend.title = element_text(size = 16, face = "bold"), 
        legend.text = element_text( size = 14),
        strip.text = element_text(size = 16, family = "Fira Sans Condensed"), 
        plot.caption = element_text(size = 14))

shift_legend2 <- function(p) {
  # ...
  # to grob
  gp <- ggplotGrob(p)
  facet.panels <- grep("^panel", gp[["layout"]][["name"]])
  empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]]))
  empty.facet.panels <- facet.panels[empty.facet.panels]
  
  # establish name of empty panels
  empty.facet.panels <- gp[["layout"]][empty.facet.panels, ]
  names <- empty.facet.panels$name
  # example of names:
  #[1] "panel-3-2" "panel-3-3"
  
  # now we just need a simple call to reposition the legend
  reposition_legend(p, 'center', panel=names)
}

shift_legend2(gr)
```

There is **important heterogeneity betweeen schools**.

We use **K-means** analysis to identify **clusters of schools according to their difference between predicted and observed attendance distribution**. We find **two main clusters**, where one of them incentivizes the **exclusion of lower-performers**. Those schools are more vulnerable and have overall lower performance.

In terms of **imputation**:

- **Overall imputation** to match school population **increases** disparities
- **Imputation to match predicted distribution** is **inbetween** no imputation and imputation for all.

```{r imputation, fig.height=5, fig.width=9, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message=FALSE}
dir_data2 = "C:/Users/mc72574/Dropbox/UT/UT Research/Asistencia/"

impute <- read.csv(paste0(dir_data2,"data/working/impute.csv"))

ggplot(data = impute, aes(x = factor(imputation, levels = c("Impute All",
                                                            "Impute Pred-Obs (All)",
                                                            "Impute Pred-Obs (Decile)",
                                                            "No Imputation")), 
                          y = mean, color = factor(cluster), fill = factor(cluster))) +
  geom_segment(aes(x = "Impute All", xend = "Impute All", y = impute$mean[impute$cluster==1 & impute$imputation=="Impute All"],
                   yend = impute$mean[impute$cluster==2 & impute$imputation=="Impute All"]), color = "grey") +
  geom_segment(aes(x = "Impute Pred-Obs (All)", xend = "Impute Pred-Obs (All)", y = impute$mean[impute$cluster==1 & impute$imputation=="Impute Pred-Obs (All)"],
                   yend = impute$mean[impute$cluster==2 & impute$imputation=="Impute Pred-Obs (All)"]), color = "grey") +
  geom_segment(aes(x = "Impute Pred-Obs (Decile)", xend = "Impute Pred-Obs (Decile)", y = impute$mean[impute$cluster==1 & impute$imputation=="Impute Pred-Obs (Decile)"],
                   yend = impute$mean[impute$cluster==2 & impute$imputation=="Impute Pred-Obs (Decile)"]), color = "grey") +
  geom_segment(aes(x = "No Imputation", xend = "No Imputation", y = impute$mean[impute$cluster==1 & impute$imputation=="No Imputation"],
                   yend = impute$mean[impute$cluster==2 & impute$imputation=="No Imputation"]), color = "grey") +
  
  geom_point(pch = 21, size = 4, stroke = 1.2) +

  scale_color_manual(name = "Cluster", labels = c("Similar","Lower att"),values = c(cols[2],cols[3])) +
  scale_fill_manual(name = "Cluster", labels = c("Similar","Lower att"),values = c(alpha(cols[2],0.8),alpha(cols[3],0.8))) +
  ylab("SIMCE Math") + xlab("") +
  ylim(180,280) +
  coord_flip()+
  theme_ipsum_fsc() + #plain
  theme(plot.margin=unit(c(0.5,0.5,0.5,0.5),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=16),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_text(size=14),
        axis.title.y = element_text(size=16),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_text(size=14),legend.position=c(0.25,0.9),
        legend.title = element_text(size = 16),
        legend.text = element_text(size=14),
        legend.background = element_rect(fill="white",colour ="grey"),
        title = element_text(size=16))


```

<!--
dir_poster <- "C:/Users/mc72574/Dropbox/Hugo/Sites/presentations/content/presentations/Attendance/ACIC_20220524/"
pagedown::chrome_print(paste0(dir_poster,"mbennett_attendance_poster.html"))
-->