---
title: "Discussion: \"Optimal tradeoffs in matched designs comparing<br/>US-trained and internationally-trained surgeons\""
subtitle: "Magdalena Bennett<br/> McCombs School of Business, UT Austin"
author: "Online Causal Inference Seminar<br/>June 29, 2021"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "custom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "macros.js"
    includes:
      in_header: header.html
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.showtext = TRUE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

theme_xaringan(
  text_color = "#333f48",
  background_color = "#FFFFFF",
  accent_color = "#900DA4",
  text_font = "Fira Mono",
  text_font_use_google = TRUE,
  title_font = "Yanone Kaffeesatz",
  title_font_use_google = TRUE
)

style_mono_accent(
  #base_color = "#bf5700",
  extra_fonts = list(google_font("Fira Sans","200","300","400","500","600"),
                     google_font("Fira Sans Condensed","200","300","400","500","600")),
  base_color = "#333f48",
  header_font_google = google_font("Yanone Kaffeesatz","200","300","400","500","600","700"),
  text_font_google   = google_font("Roboto Condensed", "300", "300i","400","500"),
  code_font_google   = google_font("Fira Mono"),
  text_bold_color = "#333f48",
  text_font_size = "125%",
  colors = c(
    lightgrey = "#C0C0C0",
    red = "#f34213",
    purple = "#900DA4",
    darkpurple = "#61077a",
    darkorange = "#db5f12",
    orange = "#ff8811",
    green = "#136f63",
    white = "#FFFFFF"),
  extra_css = list(
    ".remark-slide table" = list("display" = "table",
                   "width" = "80%",
                   "text-align" = "left"),
    ".remark-slide-number" = list("display" = "none"),
    ".strong" = list("font-weight" = "400"),
    ".big" = list("font-size" = "350%",
                     "font-family" = "Yanone Kaffeesatz",
                     "font-weight"="400"),
    ".small" = list("font-size" = "80%"),
    ".tiny" = list("font-size" = "50%"),
    ".large" = list("font-size" = "150%"),
    ".source" = list("color" = "#8c8c8c",
                     "font-size" = "80%"),
    ".remark-slide table td#highlight" = list("background-color" = "#eee1f0",
                                  "color" = "#900DA4",
                                  "font-weight" = "500"),
   # ".remark-slide table thead th" = list(),
    ".title-slide h1" = list("font-weight" = "500"),
    ".title-slide h2" = list("font-weight" = "400",
                             "font-size" =  "170%"),
    ".title-slide h3" = list("font-family" = "Roboto COndensed",
                             "font-size" = "100%",
                             "font-weight" = "200"),
    ".center2" = list("margin" = "0",
                      "position" = "absolute",
                      "top" = "50%",
                      "left" = "50%",
                      "-ms-transform" = "translate(-50%, -50%)",
                      "transform" = "translate(-50%, -50%)"),
   ".bottom2" = list("margin" = "0",
                      "position" = "absolute",
                      "top" = "90%",
                      "left" = "10%",
                      "-ms-transform" = "translate(-10%, -90%)",
                      "transform" = "translate(-10%, -90%)"),
    ".section-title h1" = list("color" = "#FFFFFF",
                               "font-size" = "2.3em",
                               "line-height" = "1.3"),
    ".medium" = list("font-size" = "1.4em"),
    ".sp-after-half" = list("margin-bottom" = "0.7em !important"),
    ".box-1,.box-1a,.box-1s,.box-1b,.box-1l,.box-1LA,.section-title-1" = list("background-color" = "#0D0887",
                                                                              "font-family" = "Yanone Kaffeesatz"),
    ".box-2,.box-2a,.box-2s,.box-2b,.box-2l,.box-2LA,.section-title-2" = list("background-color" = "#5601A4",
                                                                              "font-family" = "Yanone Kaffeesatz"),
    ".box-3,.box-3a,.box-3s,.box-3b,.box-3l,.box-3LA,.section-title-3" = list("background-color" = "#900DA4",
                                                                              "font-family" = "Yanone Kaffeesatz"),
    ".box-4,.box-4a,.box-4s,.box-4b,.box-4l,.box-4LA,.section-title-4" = list("background-color" = "#BF3984",
                                                                              "font-family" = "Yanone Kaffeesatz"),
    ".box-5,.box-5a,.box-5s,.box-5b,.box-5l,.box-5LA,.section-title-5" = list("background-color" = "#E16462",
                                                                              "font-family" = "Yanone Kaffeesatz"),
    ".box-6,.box-6a,.box-6s,.box-6b,.box-6l,.box-6LA,.section-title-6" = list("background-color" = "#db5f12",
                                                                              "font-family" = "Yanone Kaffeesatz"),
    ".box-7,.box-7a,.box-7s,.box-7b,.box-7l,.box-7LA,.section-title-7" = list("background-color" = "#FCCE25",
                                                                              "font-family" = "Yanone Kaffeesatz"),
   
   ".box-1t,.box-1tL,.section-title-1t" = list("background-color" = "rgba(13, 8, 135,0.3)",
                                      "color"="rgba(13, 8, 135,1)",
                                      "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-2t,.box-2tL,.section-title-2t" = list("background-color" = "rgba(86, 1, 164,0.3)",
                                       "color" = "rgba(86, 1, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-3t,.box-3tL,.section-title-3t" = list("background-color" = "rgba(144, 13, 164,0.3)",
                                       "color" = "rgba(144, 13, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-4t,.box-4tL,.section-title-4t" = list("background-color" = "rgba(191, 57, 132,0.3)",
                                       "color" = "rgba(191, 57, 132,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-5t,.box-5tL,.section-title-5t" = list("background-color" = "rgba(225, 100, 98,0.3)",
                                       "color" = "rgba(225, 100, 98,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-6t,.box-6tL,.section-title-6t" = list("background-color" = "rgba(248, 148, 65,0.3)",
                                       "color" = "rgba(248, 148, 65,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-7t,.box-7tL,.section-title-7t" = list("background-color" = "rgba(252, 206, 37,0.3)",
                                       "color" = "rgba(252, 206, 37,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
   
   ".box-7t, .box-6t, .box-5t, .box-4t, .box-3t, .box-2t, .box-1t" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "border-radius" = "15px"),
   
   ".box-7tL, .box-6tL, .box-5tL, .box-4tL, .box-3tL, .box-2tL, .box-1tL" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "50px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "border-radius" = "15px"),
   
    ".box-7, .box-6, .box-5, .box-4, .box-3, .box-2, .box-1" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "38px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "border-radius" = "15px"),
    ".box-7a, .box-6a, .box-5a, .box-4a, .box-3a, .box-2a, .box-1a" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
   ".box-7s, .box-6s, .box-5s, .box-4s, .box-3s, .box-2s, .box-1s" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.2em 0.2em",
                                                                      "font-weight" = "400",
                                                                      "font-size" = "100%",
                                                                      "display" = "inline",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
       ".box-7b, .box-6b, .box-5b, .box-4b, .box-3b, .box-2b, .box-1b" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "left",
                                                                      "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
   ".box-7l, .box-6l, .box-5l, .box-4l, .box-3l, .box-2l, .box-1l" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "45px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
   ".box-7LA, .box-6LA, .box-5LA, .box-4LA, .box-3LA, .box-2LA, .box-1LA" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "55px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
   ".image-80 img" = list("scale" = "80%"),
   ".pull-left-little_l" = list("float" = "left",
                                "width" = "67%"),
   ".pull-right-little_l" = list("float" = "right",
                                "width" = "27%"),
   ".pull-left-little_r" = list("float" = "left",
                                "width" = "27%"),
   ".pull-right-little_r" = list("float" = "right",
                                "width" = "67%")


  )
)

#,"li" = list("font-size" = "150%")
#    "li" = list("font-size" = "110%"),
#    "ul" = list("font-size" = "110%"),
#color palette
#5601A4
#900DA4
#F89441
#FCCE25

knitr::opts_chunk$set(message = FALSE)

```
```{css, echo = FALSE}

.small .remark-code { /*Change made here*/
  font-size: 80% !important;
}

.tiny .remark-code { /*Change made here*/
  font-size: 90% !important;
}
```

```{r setup2, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(showtext)
library(xaringanExtra)

xaringanExtra::use_scribble()

htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```

```{r fonts, message=FALSE, echo=FALSE}
font.add.google("Fira Sans Condensed", "Fira Sans Condensed")
font.add.google("Fira Sans", "Fira Sans")
```

# How do we match?

.pull-left[
- **.darkorange[Several goals are in play]**. E.g.:

  1) Balance covariate distributions.
  
  2) Minimizing distance within pairs.
  
  3) Maximizing sample size.

<br>
<br>
.box-6[How can we optimize two* of these goals at the same time?]

]

.pull-right[

]

---
# Optimal tradeoffs in matching 

- Use of **.darkorange[Pareto efficiency]** in matching:

  - Incorporate two goals in the objective function.
  
  - "Can't improve on one goal without making the other one worse".
  
  - Find a "boundary" of matched sets that vary the tradeoff between objectives.
  
--

- **.darkorange[Iterative search]** for Pareto optimal solutions:

  - Use structure of optimal sets to place informative bounds.
  
  - Depending on the multi-objective function, computing times can be high, but feasible.
  
---
# Optimal tradeoffs in matching (cont.)
  
---
# Contributions to the literature

- Maximizing sample size s.t. balancing constraints:

  - E.g. cardinality matching (Zubizarreta, Paredes, & Rosenbaum, 2014)

- Optimizing pairwise distance vs. sample size:
  
  - E.g. optimal subset matching (Rosenbaum, 2012)

--

<br>

.box-6[Combine different tradeoffs in a more general framework]

---
# Comments and Questions

--
**.darkorange[1) Importance of covariates:]**

  - Depending on the context, matching closely on a set of covariates might be more important than matching closely on others.
  
---
# Comments and Questions

**.darkorange[1) Importance of covariates:]**

  - Depending on the context, matching closely on a set of covariates might be more important than matching closely on others.
  
  - Refined covariate balance can achieve this for nominal covariates (Pimentel et al. 2015).
  
  
---
# Comments and Questions

**.darkorange[1) Importance of covariates:]**

  - Depending on the context, matching closely on a set of covariates might be more important than matching closely on others.
  
  - Refined covariate balance can achieve this for nominal covariates (Pimentel et al. 2015).
  
  - Could there be an extension for other types of covariates that need to be matched on closely?
  
---
# Comments and Questions

**.darkorange[2) External validity:]**

  - Paper proposes a way to maximize sample size $\rightarrow$ external validity

    
---
# Comments and Questions

**.darkorange[2) External validity:]**

  - Paper proposes a way to maximize sample size $\rightarrow$ external validity
  
  - Could this method incorporate a target population/penalization for loss of representation of treated units?
  
    - E.g.: Instead of maximizing sample size in $f_2$, minimize distance to a target population? 
    
  
---
# Comments and Questions

**.darkorange[3) Running time:]**

  - In cases where we are optimizing a tradeoff between balance and sample size, it could be that a projection of a MIP matching optimization problem could provide faster computational times. 

  $$\min_{\mathbf{x}} f_1(\mathbf{x}) + \rho f_2(\mathbf{x}) \ \ \rightarrow \ \ \min_{\mathbf{m}} g_1(\mathbf{m}) + \rho g_2(\mathbf{m})$$
  
---
# Comments and Questions

**.darkorange[4) Differences in estimated effects:]**

  - Matching separates the adjustment procedure and the estimation process.
  
  - Ultimately, researches chooses solution based on information of the Pareto optimal set.

  
---
# Comments and Questions

**.darkorange[4) Differences in estimated effects:]**

  - Matching separates the adjustment procedure and the estimation process.
  
  - Ultimately, researches chooses solution based on information of the Pareto optimal set.
  
  - How do results vary along the Pareto optimal set? (sensitivity to researcher's discretion)
  
---
# Final remarks

- Very interesting paper with a particularly useful result.

--

- Intuitive way of thinking about tradeoffs (that we sometimes try to solve sub-optimally).

--

- Helps make decisions in matching more transparent.

---

background-position: 50% 50%
class: center,middle

#Discussion: "Optimal tradeoffs in matched designs comparing<br/>US-trained and internationally-trained surgeons"

###Magdalena Bennett
