---
title: "How far is too far? <br/>Generalization of a regression discontinuity design away from the cutoff"
subtitle: "Magdalena Bennett"
author: "International Methods Colloquium<br/>Mar 12, 2021"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "style.css"]
    lib_dir: libs
    nature:
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
      highlightStyle: github
      highlightLines: true
      ratio: "16:9"
      beforeInit: "macros.js"
    includes:
      in_header: header.html
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.showtext = TRUE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

theme_xaringan(
  text_color = "#333f48",
  background_color = "#FFFFFF",
  accent_color = "#900DA4",
  text_font = "Fira Mono",
  text_font_use_google = TRUE,
  title_font = "Fira Sans Condensed",
  title_font_use_google = TRUE
)

style_mono_accent(
  #base_color = "#bf5700",
  extra_fonts = list(google_font("Fira Sans","200","300","400","500","600"),
                     google_font("Fira Sans Condensed"),
                     google_font("Jost")),
  base_color = "#333f48",
  header_font_google = google_font("Yanone Kaffeesatz","200","300","400","500","600","700"),
  text_font_google   = google_font("Jost"),
  code_font_google   = google_font("Fira Mono"),
  text_bold_color = "#333f48",
  text_font_size = "130%",
  colors = c(
    red = "#f34213",
    purple = "#900DA4",
    orange = "#ff8811",
    green = "#136f63",
    white = "#FFFFFF"),
  extra_css = list(
    ".body" = list("font-family" = "Jost"),
    ".remark-slide table" = list("display" = "table",
                   "width" = "80%",
                   "text-align" = "left"),
    
    #".remark-slide-number" = list("display" = "none"),
    
    ".strong" = list("font-weight" = "400"),
    ".big" = list("font-size" = "350%",
                     "font-family" = "Yanone Kaffeesatz",
                     "font-weight"="400"),
    ".small" = list("font-size" = "80%"),
    ".tiny" = list("font-size" = "50%"),
    ".tinylist" = list("font-size" = "62.5%"),
    ".source" = list("color" = "#8c8c8c",
                     "font-size" = "80%"),
    ".remark-slide table td#highlight" = list("background-color" = "#eee1f0",
                                  "color" = "#900DA4",
                                  "font-weight" = "500"),
   # ".remark-slide table thead th" = list(),
    ".title-slide h1" = list("font-weight" = "500"),
    ".title-slide h2" = list("font-weight" = "400",
                             "font-size" =  "170%"),
    ".title-slide h3" = list("font-family" = "Jost",
                             "font-size" = "100%",
                             "font-weight" = "200"),
    ".center2" = list("margin" = "0",
                      "position" = "absolute",
                      "top" = "50%",
                      "left" = "50%",
                      "-ms-transform" = "translate(-50%, -50%)",
                      "transform" = "translate(-50%, -50%)"),
    ".section-title h1" = list("color" = "#FFFFFF",
                               "font-size" = "2.3em",
                               "line-height" = "1.3"),
    ".medium" = list("font-size" = "1.4em"),
    ".sp-after-half" = list("margin-bottom" = "0.7em !important"),
    ".box-1,.box-1a,.box-1b,.section-title-1" = list("background-color" = "#0D0887"),
    ".box-2,.box-2a,.box-2b,.section-title-2" = list("background-color" = "#5601A4"),
    ".box-3,.box-3a,.box-3b,.section-title-3" = list("background-color" = "#900DA4"),
    ".box-4,.box-4a,.box-4b,.section-title-4" = list("background-color" = "#BF3984"),
    ".box-5,.box-5a,.box-5b,.section-title-5" = list("background-color" = "#E16462"),
    ".box-6,.box-6a,.box-6b,.section-title-6" = list("background-color" = "#F89441"),
    ".box-7,.box-7a,.box-7b,.section-title-7" = list("background-color" = "#FCCE25"),
    ".box-7, .box-6, .box-5, .box-4, .box-3, .box-2, .box-1" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "font-family" = "Fira Sans"),
    ".box-7a, .box-6a, .box-5a, .box-4a, .box-3a, .box-2a, .box-1a" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Fira Sans"),
       ".box-7b, .box-6b, .box-5b, .box-4b, .box-3b, .box-2b, .box-1b" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "left",
                                                                      "font-family" = "Fira Sans"),
       ".medium" = list("font-size" = "1.4em"),
    ".sp-after-half" = list("margin-bottom" = "0.7em !important"),
    ".box-1,.box-1a,.box-1s,.box-1b,.box-1l,.box-1LA,.section-title-1" = list("background-color" = "#0D0887"),
    ".box-2,.box-2a,.box-2s,.box-2b,.box-2l,.box-2LA,.section-title-2" = list("background-color" = "#5601A4"),
    ".box-3,.box-3a,.box-3s,.box-3b,.box-3l,.box-3LA,.section-title-3" = list("background-color" = "#900DA4"),
    ".box-4,.box-4a,.box-4s,.box-4b,.box-4l,.box-4LA,.section-title-4" = list("background-color" = "#BF3984"),
    ".box-5,.box-5a,.box-5s,.box-5b,.box-5l,.box-5LA,.section-title-5" = list("background-color" = "#E16462"),
    ".box-6,.box-6a,.box-6s,.box-6b,.box-6l,.box-6LA,.section-title-6" = list("background-color" = "#F89441"),
    ".box-7,.box-7a,.box-7s,.box-7b,.box-7l,.box-7LA,.section-title-7" = list("background-color" = "#FCCE25"),
   
   ".box-1t,.box-1tL,.section-title-1t" = list("background-color" = "rgba(13, 8, 135,0.3)",
                                      "color"="rgba(13, 8, 135,1)",
                                      "font-family" = "Yanone Kaffeesatz"),
    ".box-2t,.box-2tL,.section-title-2t" = list("background-color" = "rgba(86, 1, 164,0.3)",
                                       "color" = "rgba(86, 1, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-3t,.box-3tL,.section-title-3t" = list("background-color" = "rgba(144, 13, 164,0.3)",
                                       "color" = "rgba(144, 13, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-4t,.box-4tL,.section-title-4t" = list("background-color" = "rgba(191, 57, 132,0.3)",
                                       "color" = "rgba(191, 57, 132,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-5t,.box-5tL,.section-title-5t" = list("background-color" = "rgba(225, 100, 98,0.3)",
                                       "color" = "rgba(225, 100, 98,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-6t,.box-6tL,.section-title-6t" = list("background-color" = "rgba(248, 148, 65,0.3)",
                                       "color" = "rgba(248, 148, 65,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-7t,.box-7tL,.section-title-7t" = list("background-color" = "rgba(252, 206, 37,0.3)",
                                       "color" = "rgba(252, 206, 37,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
   
   ".box-7t, .box-6t, .box-5t, .box-4t, .box-3t, .box-2t, .box-1t" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center"),
   
   ".box-7tL, .box-6tL, .box-5tL, .box-4tL, .box-3tL, .box-2tL, .box-1tL" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "50px",
                                                                    "display" = "table",
                                                                    "text-align" = "center"),
   
    ".box-7, .box-6, .box-5, .box-4, .box-3, .box-2, .box-1" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "font-family" = "Yanone Kaffeesatz"),
    ".box-7a, .box-6a, .box-5a, .box-4a, .box-3a, .box-2a, .box-1a" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz"),
   ".box-7s, .box-6s, .box-5s, .box-4s, .box-3s, .box-2s, .box-1s" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.2em 0.2em",
                                                                      "font-weight" = "400",
                                                                      "font-size" = "100%",
                                                                      "display" = "inline",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz"),
       ".box-7b, .box-6b, .box-5b, .box-4b, .box-3b, .box-2b, .box-1b" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "left",
                                                                      "font-family" = "Yanone Kaffeesatz"),
   ".box-7l, .box-6l, .box-5l, .box-4l, .box-3l, .box-2l, .box-1l" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "45px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz"),
   ".box-7LA, .box-6LA, .box-5LA, .box-4LA, .box-3LA, .box-2LA, .box-1LA" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "55px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz"),
   ".image-80 img" = list("scale" = "80%"),
   ".pull-left-little_l" = list("float" = "left",
                                "width" = "67%"),
   ".pull-right-little_l" = list("float" = "right",
                                "width" = "27%"),
   ".pull-left-little_r" = list("float" = "left",
                                "width" = "27%"),
   ".pull-right-little_r" = list("float" = "right",
                                "width" = "67%")

  )
)

#,"li" = list("font-size" = "150%")
#    "li" = list("font-size" = "110%"),
#    "ul" = list("font-size" = "110%"),
#color palette
#5601A4
#900DA4
#F89441
#FCCE25
```

```{r setup2, echo=FALSE, message=FALSE, warning = FALSE}
library(knitr)
library(showtext)
library(xaringanExtra)

xaringanExtra::use_scribble()

```

```{r fonts, message=FALSE, echo=FALSE, warning = FALSE}
font_add_google("Fira Sans Condensed", "Fira Sans Condensed")
font_add_google("Fira Sans", "Fira Sans")
font_add_google("Jost", "Jost")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(ggplot2)
library(firasans)
library(hrbrthemes)
library(extrafont)
library(RColorBrewer)
library(viridis)
library(firasans)
library(hrbrthemes)
library(extrafont)
library(gridExtra)
library(pBrackets)
library(patchwork)

options(scipen = 0)
```

# Regression discontinuity design

```{r rd_intro1, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo = FALSE, warning = FALSE, message = FALSE}

set.seed(110)

n = 500

r = rnorm(n, 10,7)

treat = as.numeric(r>=10)

error = rnorm(n,0,7)

y = 2 - 1.5*(10-r) + 18*treat + 0.5*treat*(10-r) -  0.01*treat*(10-r)^2 + error

d = data.frame("r" = r,
               "treat" = treat,
               "y" = y)

ggplot(data = d, aes(x=r, y=y)) +
  geom_point(size = 3, pch = 21, color = "dark grey", fill="white") +
  geom_vline(aes(xintercept=10), lty = 2, color = "#333f48", lwd = 1.5) +
  ggplot2::annotate("rect", color="white", fill="white", xmin = 7, xmax = 13, ymin=38, ymax = 48) +
  ggplot2::annotate("text", label = "Cutoff", size = 10, hjust = 0.5, x = 10, y = 43, color = "#333f48",
                    family = "Fira Sans Condensed")+
  theme_bw() +
  theme_ipsum_fsc()+
  xlab("Running variable") + ylab("Outcome") + 
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(colour = "light grey"),
        panel.grid.minor.y = element_blank(),
        axis.line = element_blank())+
  theme(axis.title.x = element_text(size = 24, color = "#333f48"),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 24, color = "#333f48"),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_blank(),legend.position="none",
        legend.title = element_text(size=12),
        legend.text = element_text(size=10),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=12))


```

---
# Increasingly popular method


```{r rd_publications, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo = FALSE, warning = FALSE, message = FALSE}
pub <- read.csv("https://raw.githubusercontent.com/maibennett/presentations/main/content/presentations/RD/IC_20210312/data/regression_discontinuityv2.csv")

names(pub) <- c("year","results")

ggplot(data = pub, aes(x=year, y=results)) +
  geom_point(size = 4, pch = 21, color = "#E16462", fill="#E16462") +
  geom_line(color = "#E16462", lty = 2, lwd = 1.5) +
  theme_bw() +
  theme_ipsum_fsc()+
  xlab("Year") + ylab("Number of publications") + 
  labs(caption = "Note: Publications including 'Regression Discontinuity' in their title or abstract for research fields of Social Sciences.\nSource: Dimensions.ai") + 
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(colour = "light grey"),
        panel.grid.minor.y = element_blank(),
        axis.line = element_blank())+
  theme(axis.title.x = element_text(size = 24, color = "#333f48"),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 24, color = "#333f48"),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_text(size = 14),legend.position="none",
        legend.title = element_text(size=12),
        legend.text = element_text(size=10),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=28, color = "#333f48"))
```
---
# Strong internal validity

```{r rd_intro2, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo = FALSE, warning = FALSE, message = FALSE}

ggplot(data = d, aes(x=r, y=y)) +
  geom_point(size = 3, pch = 21, color = alpha("dark grey",0.8), fill="white") +
  geom_vline(aes(xintercept=10), lty = 2, color = "grey", lwd = 1.5) +
  
  geom_smooth(data = dplyr::filter(d, r<10), method = "gam", se = FALSE, color = "grey", lwd = 2) + 
  geom_smooth(data = dplyr::filter(d, r>=10), method = "gam", se = FALSE, color = "grey", lwd = 2) + 
  
  ggplot2::annotate("rect", color="white", fill="white", xmin = 11, xmax = 14, ymin=11, ymax = 15) +
  ggplot2::annotate("text", label = "LATE", size = 10, hjust = 0.5, x = 12.5, y =13, color = "#E16462",
                    family = "Fira Sans Condensed")+
  
  ggplot2::annotate("segment",x = 10, xend = 10, y = 1.44, yend = 19.6, color = "#E16462", lwd = 2) +
  
  theme_bw() +
  theme_ipsum_fsc()+
  xlab("Running variable") + ylab("Outcome") + 
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(colour = "light grey"),
        panel.grid.minor.y = element_blank(),
        axis.line = element_blank())+
  theme(axis.title.x = element_text(size = 24, color = "#333f48"),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 24, color = "#333f48"),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_blank(),legend.position="none",
        legend.title = element_text(size=12),
        legend.text = element_text(size=10),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=12))


```

---
# ... But limited external validity

```{r rd_intro3, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo = FALSE, warning = FALSE, message = FALSE}

ggplot(data = d, aes(x=r, y=y)) +
  geom_point(size = 3, pch = 21, color = alpha("dark grey",0.8), fill="white") +
  geom_vline(aes(xintercept=10), lty = 2, color = "grey", lwd = 1.5) +
  
  geom_smooth(data = dplyr::filter(d, r<10), method = "gam", se = FALSE, color = "#E16462", lwd = 2) + 
  geom_smooth(data = dplyr::filter(d, r>=10), method = "gam", se = FALSE, color = "#E16462", lwd = 2) + 
  
  ggplot2::annotate("rect", color="white", fill="white", xmin = 11, xmax = 14, ymin=11, ymax = 15) +
  ggplot2::annotate("text", label = "LATE", size = 10, hjust = 0.5, x = 12.5, y =13, color = "grey",
                    family = "Fira Sans Condensed")+
  
  ggplot2::annotate("segment",x = 10, xend = 10, y = 1.44, yend = 19.6, color = "grey", lwd = 2) +
  
  theme_bw() +
  theme_ipsum_fsc()+
  xlab("Running variable") + ylab("Outcome") + 
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(colour = "light grey"),
        panel.grid.minor.y = element_blank(),
        axis.line = element_blank())+
  theme(axis.title.x = element_text(size = 24, color = "#333f48"),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 24, color = "#333f48"),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_blank(),legend.position="none",
        legend.title = element_text(size=12),
        legend.text = element_text(size=10),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=12))


```

---
# Missing data and no overlap

```{r rd_intro4, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo = FALSE, warning = FALSE, message = FALSE}

y1 = 2 - 1.5*(10-r) + 18 + 0.5*(10-r) -  0.01*(10-r)^2 + error
y0 = 2 - 1.5*(10-r) + error

d$y1 <- y1
d$y0 <- y0

ggplot(data = d, aes(x=r, y=y)) +
  geom_point(size = 3, pch = 21, color = alpha("dark grey",0.8), fill="white") +
  geom_vline(aes(xintercept=10), lty = 2, color = "grey", lwd = 1.5) +
  
  geom_smooth(data = dplyr::filter(d, r<10), method = "gam", se = FALSE, color = "#E16462", lwd = 2) + 
  geom_smooth(data = dplyr::filter(d, r>=10), method = "gam", se = FALSE, color = "#E16462", lwd = 2) +
  
  geom_smooth(data = dplyr::filter(d, r<10), aes(x = r, y = y1), method = "gam", se = FALSE, color = "#E16462", lwd = 2, lty = 2) + 
  geom_smooth(data = dplyr::filter(d, r>=10), aes(x = r, y = y0), method = "gam", se = FALSE, color = "#E16462", lwd = 2, lty = 2) +
  
  ggplot2::annotate("rect", color="white", fill="white", xmin = 11, xmax = 14, ymin=11, ymax = 15) +
  ggplot2::annotate("text", label = "LATE", size = 10, hjust = 0.5, x = 12.5, y =13, color = "grey",
                    family = "Fira Sans Condensed")+
  
  ggplot2::annotate("segment",x = 10, xend = 10, y = 1.44, yend = 19.6, color = "grey", lwd = 2) +
  
  theme_bw() +
  theme_ipsum_fsc()+
  xlab("Running variable") + ylab("Outcome") + 
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(colour = "light grey"),
        panel.grid.minor.y = element_blank(),
        axis.line = element_blank())+
  theme(axis.title.x = element_text(size = 24, color = "#333f48"),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 24, color = "#333f48"),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_blank(),legend.position="none",
        legend.title = element_text(size=12),
        legend.text = element_text(size=10),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=12))


```


---
# Can we find a generalization interval?

```{r rd_intro5, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo = FALSE, warning = FALSE, message = FALSE}

ggplot(data = d, aes(x=r, y=y)) +
  geom_point(size = 3, pch = 21, color = alpha("dark grey",0.8), fill="white") +
  geom_vline(aes(xintercept=10), lty = 2, color = "grey", lwd = 1.5) +
  
  geom_smooth(data = dplyr::filter(d, r<10), method = "gam", se = FALSE, color = "grey", lwd = 2) + 
  geom_smooth(data = dplyr::filter(d, r>=10), method = "gam", se = FALSE, color = "grey", lwd = 2) + 
  
  geom_vline(aes(xintercept = 5), lty = 2, color = "#E16462", lwd = 2) +
  geom_vline(aes(xintercept = 15), lty = 2, color = "#E16462", lwd = 2) +
  
  ggplot2::annotate("rect", color="white", fill="white", xmin = 11, xmax = 14, ymin=11, ymax = 15) +
  ggplot2::annotate("text", label = "LATE", size = 10, hjust = 0.5, x = 12.5, y =13, color = "grey",
                    family = "Fira Sans Condensed")+
  
  ggplot2::annotate("segment",x = 10, xend = 10, y = 1.44, yend = 19.6, color = "grey", lwd = 2) +
  
  theme_bw() +
  theme_ipsum_fsc()+
  xlab("Running variable") + ylab("Outcome") + 
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(colour = "light grey"),
        panel.grid.minor.y = element_blank(),
        axis.line = element_blank())+
  theme(axis.title.x = element_text(size = 24, color = "#333f48"),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 24, color = "#333f48"),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_blank(),legend.position="none",
        legend.title = element_text(size=12),
        legend.text = element_text(size=10),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=12))

```

---
# This paper

**Identification of generalization interval and estimation of ATT for population within such interval:**
--

  - Pre-intervention period informs the generalization bandwidth <br>
  .tiny[(Wing & Cook, 2013; Keele, Small, Hsu, & Fogarty, 2020)]
    
  - Leverage the use of predictive covariates for breaking link between running variable and outcome <br>
  .tiny[(Angrist & Rokkanen, 2015; Rokkanen, 2015; Keele, Titiunik, & Zubizarreta, 2015)]
  
  - Based on idea of local randomization near the cutoff <br>
  .tiny[(Lee, 2008; Cattaneo, Frandsen, & Titiunik, 2015)]
  
---
# This paper

**Main advantages:**
--

- Gradual approach
  - No need for "All or Nothing"
  - Interval informed by the data .tinylist[(Cattaneo et al., 2015)]
--

- No extrapolation of population characteristics
  - Compare like-to-like .tinylist[(Rosenbaum, 1987)]
  - Makes overlap region explicit
--

- Generalization to population of interest
  - Use of representative template matching .tinylist[(Silber et al., 2014; Bennett, Vielma, & Zubizarreta, 2020)]
--

- Sensitivity analysis to hidden bias .tiny[(Rosenbaum, 2010; Keele et al., 2020)]

---
# Outline

1. Motivation
<br>
<br>
2. Generalized Regression Discontinuity Design (GRD)

  2.1 Framework
  
  2.2 GRD in practice
<br>
<br>
3. Application: Free Higher Education in Chile
<br>
<br>
4. Conclusions

---

background-position: 50% 50%
class: left, bottom, inverse, nonum
.big[
Get me away from here...
]

---
# Generalized Regression Discontinuity Design (GRD)

--
Two-part problem with pre- and post-intervention periods:
<br>

--
<br>
<span class="box-5">1) Identification of generalization interval H<sup>\*</sup></span>
<br>
.box-5t[(using pre-intervention period)]

--
<br>
<span class="box-5">2) Estimation of ATT for population within H<sup>\*</sup></span>
<br>
.box-5t[(using post-intervention period)]

---
#The setup

- **Two periods**: pre- and post-intervention, $t=0$ and $t=1$.

- Running variable $R$ determines assingment $Z$ in $t=1$. E.g.:
$$Z_{it} = \mathrm{I}(R_{it}<c)$$

- Potential outcomes under treatmet $z = 0,1$:
$$Y^{(z)}_{it} = g_z(\mathbf{X}_{it},\mathbf{u}_{it},r_{it}) + z_{it}\cdot \underbrace{\tau(\mathbf{X}_{it},\mathbf{u}_{it},r_{it})}_{\color{#E16462}{\text{Treat. Effect}}} + \underbrace{\alpha_t}_{\color{#E16462}{\text{Period FE}}}$$
  - $\mathbf{X}_{it}$: Predictive covariates
  
  - $\mathbf{u}_{it}$: Unobserved confounders
  
  - $\tau(\cdot)$: Causal effect
  
---
# Two periods for GRD

```{r grd_setup, fig.height=5.7, fig.width=12, fig.align='center', dev='svg', echo = FALSE, warning = FALSE, message = FALSE}
set.seed(100)

df_pre <- data.frame(xaxisTime=runif(300)*20) %>%
  mutate(Y = .3*xaxisTime+rnorm(300),
         state="1",
         groupX=floor(xaxisTime)+.5,
         groupLine=floor(xaxisTime),
         cutLine=rep(c(9,11),150)) %>%
  group_by(groupX) %>%
  mutate(mean_Y=mean(Y)) %>%
  ungroup() %>%
  arrange(groupX)

df_pre$id = seq(1,nrow(df_pre),1)

df_post <- data.frame(xaxisTime=runif(300)*20) %>%
  mutate(Y = .3*xaxisTime+5*(xaxisTime>10)-.15*xaxisTime*(xaxisTime>10)+rnorm(300),
         state="1",
         groupX=floor(xaxisTime)+.5,
         groupLine=floor(xaxisTime),
         cutLine=rep(c(9,11),150)) %>%
  group_by(groupX) %>%
  mutate(mean_Y=mean(Y)) %>%
  ungroup() %>%
  arrange(groupX)

y1 = df_pre$Y + exp(-2*(df_pre$xaxisTime-3)) + 1 - 0.04*df_pre$xaxisTime
poly_fit_y1 <- as.data.frame(lowess(df_pre$xaxisTime, y1, f=0.4))

df_post2 <- data.frame(xaxisTime=runif(300)*20) %>%
  mutate(Y = .3*xaxisTime+(xaxisTime<10)*(poly_fit_y1$y-df_pre$Y)+rnorm(300),
         state="1",
         groupX=floor(xaxisTime)+.5,
         groupLine=floor(xaxisTime),
         cutLine=rep(c(9,11),150)) %>%
  group_by(groupX) %>%
  mutate(mean_Y=mean(Y)) %>%
  ungroup() %>%
  arrange(groupX)

df_post2_detrend <- data.frame(xaxisTime=runif(300)*20) %>%
  mutate(Y = rnorm(300)+2.8 + (xaxisTime<10)*(poly_fit_y1$y-df_pre$Y),
         state="1",
         groupX=floor(xaxisTime)+.5,
         groupLine=floor(xaxisTime),
         cutLine=rep(c(9,11),150)) %>%
  group_by(groupX) %>%
  mutate(mean_Y=mean(Y)) %>%
  ungroup() %>%
  arrange(groupX)

df_post$id = seq(1,nrow(df_post),1)
df_post2$id = seq(1,nrow(df_post2),1)
df_post2_detrend$id = seq(1,nrow(df_post2_detrend),1)


poly_fit_y0 <- as.data.frame(lowess(df_pre$xaxisTime, df_pre$Y, f=0.4))

y1 = df_pre$Y + exp(-2*(df_pre$xaxisTime-3)) + 1 - 0.04*df_pre$xaxisTime
poly_fit_y1 <- as.data.frame(lowess(df_pre$xaxisTime, y1, f=0.4))

#Pre-period
g1 <- ggplot(df_pre,aes(y=Y,x=xaxisTime, label=NA), fill="#333f48")+geom_point(pch=21, fill="white", color="#333f48")+
  geom_line(data = poly_fit_y0, aes(x=x, y=y, color='#E16462'), size=1)+
  geom_line(data = poly_fit_y1, aes(x=x, y=y, color = '#900DA4'), size=1, lty=2)+
  geom_vline(aes(xintercept=10),linetype='dashed') + theme_bw()+
  xlab("Running variable") + ylab("Outcome")+ggtitle("Pre-intervention")+
  scale_colour_identity(name='',breaks=c('#E16462','#900DA4'),
                        labels = c(bquote(Y^{(0)}~(R)),bquote(Y^{(1)}~(R))), guide="legend") +
  theme_bw() +
  theme_ipsum_fsc()+
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"))+
  theme(axis.title.x = element_text(size=20,margin = margin(t = 10, r = 0, b = 0, l = 0), color = "#333f48"),
        axis.text.x = element_text(size=15),
        axis.title.y = element_text(size=20,margin = margin(t = 0, r = 10, b = 0, l = 0), color = "#333f48"),
        axis.text.y = element_text(size=15),legend.position="bottom",
        title = element_text(color = "#333f48", size = 22, face = "plain"),
        legend.text = element_text(size=15, color = "#333f48"),
        legend.title = element_text(size=18, color = "#333f48"))+
  ylim(-2,10)


#Post-period
g2 <- ggplot(df_post2,aes(y=Y,x=xaxisTime, label=NA))+geom_point(pch=21, fill="white", color="#333f48")+
  geom_line(data = poly_fit_y0[poly_fit_y0$x>10,], aes(x=x, y=y,color='#E16462'), size=1)+
  geom_line(color='#E16462',data = poly_fit_y0[poly_fit_y0$x<10,], aes(x=x, y=y), size=1,lty=2)+
  geom_line(data = poly_fit_y1[poly_fit_y1$x>10,], aes(x=x, y=y, color='#900DA4'), size=1, lty=2)+
  geom_line(color='#900DA4',data = poly_fit_y1[poly_fit_y1$x<10,], aes(x=x, y=y), size=1, lty=1)+
  geom_vline(aes(xintercept=10),linetype='dashed') + theme_bw()+
  xlab("Running variable") + ylab("Outcome")+ggtitle("Post-intervention")+
  scale_colour_identity(name='',breaks=c('#E16462','#900DA4'),
                        labels = c(bquote(Y^{(0)}~(R)),bquote(Y^{(1)}~(R))), guide="legend") +
  theme_bw() +
  theme_ipsum_fsc()+
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"))+
  theme(axis.title.x = element_text(size=20,margin = margin(t = 10, r = 0, b = 0, l = 0), color = "#333f48"),
        axis.text.x = element_text(size=15),
        axis.title.y = element_text(size=20,margin = margin(t = 0, r = 10, b = 0, l = 0), color = "#333f48"),
        axis.text.y = element_text(size=15),legend.position="bottom",
        title = element_text(color = "#333f48", size = 22, face = "plain"),
        legend.text = element_text(size=15, color = "#333f48"),
        legend.title = element_text(size=18, color = "#333f48")) +
  ylim(-2,10)

g1 + g2 + plot_layout(guides = "collect") &
  theme(legend.position='bottom')
```
---
# A gradual approach

- Conditional expectations of potential outcomes, $Y^{(z)}_t(R)$:
$$Y^{(0)}_0(R) = \mathbb{E}[Y^{(0)}_{i0}|R] = \mu_0(R)$$
$$Y^{(1)}_0(R) = \mathbb{E}[Y^{(1)}_{i0}|R] = \underbrace{\mu_0(R)}_{\color{#900DA4}{\text{Avg. Outcome by R}}} + \underbrace{\tau_0(R)}_{\color{#900DA4}{\text{Treat. Effect by R}}}$$

--

- Identify generalization interval $H = [H_{-},H_{+}]$ for $t=0$:
$$R_{i} = h(\mathbf{X}_{i}) + \eta_i \ \ \forall \ R_i \in H$$
--

- If $H^* = \max\{|H|\}$ exists, then for a set of covariates $\mathbf{X} = \mathbf{X}_T$:
$$Y^{(0)}_0(R')|\mathbf{X}_T = Y^{(0)}_0(R'')|\mathbf{X}_T \ \ \text{for any} \ R', R'' \in H^*$$

---
# Conditional outcome within the generalization interval

```{r grd_setup_pre, fig.height=5.7, fig.width=12, fig.align='center', dev='svg', echo = FALSE, warning = FALSE, message = FALSE}
coef <- summary(lm(Y ~ xaxisTime, data = df_pre))$coef



df_pre_detrend <- df_pre %>% mutate(Y = Y - coef[2,1]*xaxisTime - coef[1,1]) %>%
  filter(xaxisTime>=5 & xaxisTime<=15)

#Post-period after matching
ggplot(df_pre_detrend,aes(y=Y,x=xaxisTime, label=NA), fill="slategrey")+geom_point(pch=21, fill="white", color="#333f48") + 
  geom_smooth(method = "loess")+ylim(0,10) + xlim(4,16)
  geom_line(color='deepskyblue3',data = poly_fit_y0[poly_fit_y0$x<10,], aes(x=x, y=2.8), size=1,lty=2)+
  geom_line(data = poly_fit_y1[poly_fit_y1$x>10,], aes(x=x, y=y, color='darkorchid3'), size=1, lty=2)+
  geom_line(color='darkorchid3',data = poly_fit_y1[poly_fit_y1$x<10,], aes(x=x, y=y), size=1, lty=1)+
  geom_vline(aes(xintercept=10),linetype='dashed', color="black") +
  geom_vline(aes(xintercept=5),linetype='dashed',color='springgreen4',lwd=1) +
  geom_vline(aes(xintercept=15),linetype='dashed',color='springgreen4',lwd=1) + theme_bw()+
  xlab("Running variable") + ylab("Outcome")+ggtitle("Post-intervention after Matching")+
  scale_colour_identity(name='Potential Outcomes',breaks=c('deepskyblue3','darkorchid3',"springgreen4"),
                        labels = c("Y0|X","Y1|X","Matched obs"),lty=c(1,1,NA),
                        pch=c(NA,NA,1), guide="legend") +
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"))+
  theme(axis.title.x = element_text(size=15,margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_text(size=12),
        axis.title.y = element_text(size=15,margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_text(size=12),legend.position="bottom") +
  ylim(0,10) + xlim(4,16)
```